### 2012-01-20 contents.tex
Writing Methods section.

\begin{itemize}
	\item What activity is actually performed when a bug fix is marked as verified?
	\item When is the verification performed (i.e., just after the fix, or is there a specific period for the verification of fixed bugs?)
	\item How is the distribution of verification across developers?
	\item Is there a dedicated team for QA?
\end{itemize}

@@

\begin{section}{Background}
\end{section}

  \begin{subsection}{Bug Tracking Systems}
Bug workflow. Bugzilla. \cite{Bird2009}

@@

@@

Similarly, what do developers do before marking a bug as VERIFIED? Bugzilla documentation states that, when a bug is VERIFIED, it means that ``QA [quality assurance team] has looked at the bug and the resolution and agrees that the appropriate resolution has been taken''\cite{\url{https://landfill.bugzilla.org/bugzilla-3.6-branch/page.cgi?id=fields.html}}. Again, the definition is loose, although it assumes the existence of a QA role.

The developer documentation for both Eclipse and Netbeans are unclear about how the verification of a bug fix is performed. So, what forms of  software verification are they used in the verification process of these projects?
% In some parts they mention running unit tests, and running a patched version of the software and try to reproduce the bug.

@@

@@

@@

\begin{section}{Data Extraction}

	We have had access to Eclipse Bugzilla database from October, 2001 to June, 2010, in MySQL format. The data was made available as part of the Mining Software Repositories 2011 Challenge\footnote{\url{http://www.msrconf.org/msr-challenge.html}}.

	The database includes data from 155 Eclipse subprojects. However, we have discarded subprojects in which no bug was ever reopened, and subprojects with less than 50 reported bugs.

	For the remaining 34 subprojects, we have selected only bugs that were verified until December 2009. The rationale is that if, within six months, a verified bug was not reopened, it probably will not be reopened after that and it is, thus, fully fixed.

\begin{subsection}{The Eclipse Project}
	History. \\
	Recommendations about Bugzilla usage. Four-eyes principle.

	The Eclipse Project was created in 2001 by IBM and released into open source in 2004\footnote{\url{http://www.eclipse.org/org/}}, with focus on the creation of a extensible platform that supports the software development lifecycle. Today, it is maintained by its open source community, as well as by IBM employees and people from other organizations.

	The Eclipse Project consists of many subprojects, including the Platform project and dozens of plugins. Although each subproject is maintained by a distinct team and is relatively autonomous, all of them should follow a general, documented process.

	All subprojects use Bugzilla for bug reporting, and there are clear rules regarding the process of change in the software\footnote{\url{http://wiki.eclipse.org/Development_Resources/HOWTO/Bugzilla_Use}}. Here is an excerpt of a document in the Eclipse web site that explains the bug fixing process\footnote{\url{http://wiki.eclipse.org/Bug_Reporting_FAQ}} (emphasis added):

	\begin{quote}
	When a new bug is entered it begins life with a Status of either \textbf{Unconfirmed} for normal users or \textbf{New} for users with commit privileges. The bug is typically assigned to the component owner. The component owner will usually use a query of Status = Unconfirmed or New and Assigned to = me to browse what is essentially the component's inbox. She or he will assign bug reports to developers. Email will be generated to the developer (this part is configurable). (Note that the act of reassigning a bug to somebody changes its status to New). The developer may change unconfirmed bugs to new.

	The assigned developer will “accept” the bug which will change its status to \textbf{Assigned}. After working on the bug the developer will mark the bug as \textbf{Resolved} and will select a resolution (Fixed, Invalid, Wontfix, Worksforme).

	Once a bug is resolved there are still a few states it can transition to. How you choose to use these states will be up to individual teams. In our example, when a test pass comes up a component owner can build a query that searched for resolved bugs. After testing that the fix worked a resolved bug can be transitioned to \textbf{verified}, directly to \textbf{closed}, or in fact, \textbf{reopened}. By searching for bugs with a Status of verified and a resolution of fixed developers can come up with release notes. A verified/fixed bug can then be transitioned to closed. And yes, closed bugs can be reopened if need be. As an added bonus other bug status can be transitioned through verified to closed as well. This gives developers the opportunity to test worksforme claims.
	\end{quote}

	There’s also a recommendation to follow the four eyes principle in the verification process (emphasis added):

	\begin{quote}
	When a developer fixes the bug, the status is set to RESOLVED - FIXED, and it is assigned to another developer on the team to verify. \textbf{It is important that the verifier be a different person than the fixer} because the fixer is too close to the code and thus may not be as diligent at testing the corner cases.
	\end{quote}

	Finally, there are additional guidelines regarding the workflow:

	\begin{quote}
		\begin{itemize}
	\item All bugs should be verified before the next integration build.

	\item When a committer verifies a fix, the status is changed to VERIFIED.

	\item When the project does a major release, the VERIFIED bugs are changed to CLOSED.
		\end{itemize}
	\end{quote}
\end{subsection}

\begin{subsection}{The Netbeans Project}
	History. \\
	Recommendations for developers.
\end{subsection}

\begin{section}{Data Sampling}
	Scripts to classify interesting cases.
	Random sampling stratified by case.

\begin{section}{Analysis}
	grounded theory-style.

@@

\begin{section}{Experimental Setup}

\begin{verbatim}
Data extraction
Data processing: 
- filtering (bugs up to december 20xx)
- classification script
Sampling:
- random sampling (for coding)
- entire data set
Analysis:
- grounded theory (open coding)
- latent Dirichlet allocation
- descriptive statistics
\end{verbatim}

\begin{section}{Results}

	O que quer dizer “a solução de um bug foi verificada”?

	Como já foi comentado anteriormente, o Bugzilla permite que os relatórios de bugs sejam marcados como RESOLVED (resolvido) e, depois, como VERIFIED (verificado). Mas afinal, o que significa verificar a solução de um bug? Que atividades ou checagens foram realizadas antes de se marcar o bug como VERIFIED?

	A documentação do Bugzilla tem algo a dizer sobre isso o status VERIFIED:

	“QA [quality assurance team] has looked at the bug and the resolution and agrees that the appropriate resolution has been taken.”

	Essa definição sugere que a verificação deve ser feita por uma equipe especializada em controle de qualidade. Além disso, o verificador deve concordar que a resolução é apropriada. A definição do Bugzilla para por aí, mas eu consigo imaginar alguns critérios para considerar uma resolução apropriada:

	O bug não se manifesta após a aplicação da solução.
	A solução não gera novos bugs.
	O código da solução segue as convenções de codificação e políticas do projeto.
	A solução possui uma boa relação custo-benefício — isto é, ou não existem soluções melhores ou a solução é um paliativo por conta da alta complexidade de uma solução melhor.
	Minha expectativa, portanto, é que, antes de o bug ser marcado como VERIFICADO, ele passe pelas seguintes atividades:

	Aplicação do patch com a solução do bug na versão do software onde o bug foi encontrado.
	Execução do software com o bug, seguindo os passos para reprodução do bug, e verificando que o bug não se manifesta
	Inspeção do código-fonte. Deve ser verificada a aderência às convenções de codificação do projeto e a outras políticas (por exemplo, cada patch deve vir com testes de unidade).
	Execução de testes de unidade (se houver). Deve se verificar que todos os testes passam.
	Bom, mas isso é a teoria. Quais atividades são executadas na prática, em projetos reais, antes de se marcar um bug como VERIFIED? Para responder a essa pergunta, analisei um total de 80 bugs, considerando os projetos Eclipse/EMF, Eclipse/Platform, Netbeans/versioncontrol e Netbeans/profiler. Eis alguns casos verificados nesses projetos:

	O reporter (ou outro usuário que também enfrenta o problema) verificou que, em uma build do produto que inclui a solução, o bug deixou de se manifestar. Aparentemente não há inspeção do código-fonte, e nem uma equipe de QA. Situação frequente no Netbeans. Pode ser um procedimento necessário quando os desenvolvedores não conseguem reproduzir o bug.
	A solução está disponível em um build que é divulgado no site do projeto. Esse critério é usado pelo projeto Eclipse/EMF e foi decidido em reunião. O projeto usa o status FIXED para bugs cuja solução está no CVS.
	Bugs que já foram resolvidos há muito tempo são marcados como verificados, por se entender que, se nenhuma problema ocorreu até então, isso significa que a solução é adequada.
	A mesma pessoa que resolve o bug o marca como VERIFIED logo depois. Aparentemente não existe verificação de fato nesse caso.
	O desenvolvedor que resolve o bug solicita que outro desenvolvedor verifique a solução. Comum no Eclipse/Platform.
	No bug 249436 do Eclipse/EMF, um patch é submetido por um desenvolvedor, mas outro desenvolvedor acaba reescrevendo o patch, fornecendo evidências de que o código-fonte foi inspecionado.
	No bug 269789 do Eclipse/EMF, há evidências da elaboração e da execução de testes de unidade.
	A análise sugere que o status VERIFIED significa coisas diferentes em diferentes projetos. A execução de uma versão corrigida do software parece ser o método predominante de verificação. Em alguns casos, o status VERIFIED não representa verificação, e sim a indicação de que a solução está disponível para o usuário final. Evidências de inspeção de código e execução de testes de unidade só foram encontradas no projeto Eclipse/Platform.

\begin{section}{Conclusion}
\end{section}

>>>>

\begin{itemize}
	\item How is the verification actually performed? Are there performed ad hoc tests, automated tests, code inspection?
	\item When is the verification performed? Is it performed just after the fix, or is there a specific period for the verification of fixed bugs?
	\item Who performs the verification? Is there a dedicated team for QA? 
	% \item How is the distribution of verification across developers?
	% \item Is there a dedicated team for QA?
\end{itemize}

@@

\begin{section}{Background}
% Bug workflow. Bugzilla. \cite{Bird2009}

@@

\TODO Roles: reporter, fixer, verifier, reopener, closer.

@@

Similarly, what do developers do before marking a bug as VERIFIED? Bugzilla documentation states that, when a bug is VERIFIED, it means that ``QA [quality assurance team] has looked at the bug and the resolution and agrees that the appropriate resolution has been taken''\footnote{\url{https://landfill.bugzilla.org/bugzilla-3.6-branch/page.cgi?id=fields.html}}. Again, the definition is loose, although it assumes the existence of a QA role.

The developer documentation for both Eclipse and NetBeans are unclear about how the verification of a bug fix is performed. So, what forms of  software verification are they used in the verification process of these projects?
% In some parts they mention running unit tests, and running a patched version of the software and try to reproduce the bug.

@@

% O bug não se manifesta após a aplicação da solução.
% A solução não gera novos bugs.
% O código da solução segue as convenções de codificação e políticas do projeto.
% A solução possui uma boa relação custo-benefício — isto é, ou não existem soluções melhores ou a solução é um paliativo por conta da alta complexidade de uma solução melhor.
% 
% Minha expectativa, portanto, é que, antes de o bug ser marcado como VERIFICADO, ele passe pelas seguintes atividades:
% 
% Aplicação do patch com a solução do bug na versão do software onde o bug foi encontrado.
% Execução do software com o bug, seguindo os passos para reprodução do bug, e verificando que o bug não se manifesta
% Inspeção do código-fonte. Deve ser verificada a aderência às convenções de codificação do projeto e a outras políticas (por exemplo, cada patch deve vir com testes de unidade).
% Execução de testes de unidade (se houver). Deve se verificar que todos os testes passam.

	% \item How is the verification actually performed?
	% \item When is the verification performed (i.e., just after the fix, or is there a specific period for the verification of fixed bugs?)
	% \item Who performs the verification?

In order to answer our exploratory research questions, we followed the following method:

\begin{enumerate}
	\item \textbf{Data extraction}: we have obtained publicly available raw data from the Bugzilla repositories of two popular integrated development environments, Eclipse and Bugzilla.

	\item \textbf{Data sampling}: we have chosen two subprojects. After that, we sampled bugs with interesting characteristics.

	\item \textbf{Data analysis}: we have read the full bug reported for the sampled bugs. Also, we have performed quantitative analysis on all bugs.

	\item ...
\end{enumerate}

\begin{subsection}{Available Data}

	In order to perform the desired analysis, we needed access to all data recorded by Bugzilla for a specific project. We have found such data for two projects --- Eclipse and NetBeans --- from the domain of integrated development environments. The data were made available as part of the Mining Software Repositories 2011 Challenge\footnote{\url{http://www.msrconf.org/msr-challenge.html}} in the form of MySQL database dumps. The files contain all tables from the respective databases, except for the tables containing information about developers.

	% We have had access to Eclipse Bugzilla database from October, 2001 to June, 2010, and to NetBeans Bugzilla database from ... to .... The data was made available as part of the Mining Software Repositories 2011 Challenge\footnote{\url{http://www.msrconf.org/msr-challenge.html}} as a SQL dump. The SQL dump contains all information recorded by Bugzilla for the projects (except for information about developers, which was removed for privacy reasons).

	Eclipse development began in late 1998 with IBM\footnote{\url{http://www.ibm.com/developerworks/rational/library/nov05/cernosek/}}. It was licensed as open source in November, 2001. In 2004, IBM created the Eclipse Foundation in order to reduce the perception of Eclipse as a IBM-controlled project.

	NetBeans\footnote{\url{http://netbeans.org/about/history.html}} started  as a student project in 1996, was bought by Sun Microsystems in Oct, 1999, and was open sourced in Mar, 2000.

	Both Eclipse and NetBeans are composed of many subprojects. The data comprises all bug reports from all Eclipse subprojects from October, 2001 to June, 2010, and from all NetBeans subprojects from June, 1998 to June, 2010.

	There 316,911 bug reports for 155 subprojects in Eclipse, and 185,578 bug reports for 39 subprojects in NetBeans.

@@

\begin{subsection}{Sampling of Bug Reports}

	In order to obtain insight about the research questions, we read a sample of bug reports from both projects. For each project, two subprojects were chosen randomly among the projects with most verified bugs, resulting in the following subprojects: Eclipse/Platform, Eclipse/EMF, Netbeans/profiler, Netbeans/versioncontrol. We have sampled 20 bug reports for each subproject, totalizing 80 bug reports.

\begin{subsection}{Analysis}

	Each of the sampled bugs was read.

@@

\begin{section}{Results and Discussion}
	Some insights were gained in this step:

	\begin{itemize}
		\item In some projects, the status VERIFIED does not correspond to software verification. In Eclipse/EMF, for example, it means that the fix is available in a build that is promoted on the project website\footnote{\url{http://wiki.eclipse.org/Modeling_PMC_Meeting,_2007-10-16}}.
		\item In the NetBeans subprojects that were analyzed, it was common that the reporter also verified the fix. In some bugs, this was necessary since the developer was not able to reproduce the bug.
		\item Often, the fixer verifies the bug just after marking it as fixed.
		\item Sometimes, several old bugs are marked as VERIFIED at the same time.
		\item In Eclipse/Platform, it is common that the fixer requests someone else to verify the fix.
		\item In Eclipse/EMF, bug #249436, one developer rewrites the patch submitted by other developer.
		\item In Eclipse/EMF, bug #269789, a test case is used to assess the proposed fix.
	\end{itemize}

@@
